---
title: "Comparing endid and lwdidR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing endid and lwdidR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Both `endid` and `lwdidR` implement the panel Difference-in-Differences (DiD) approach proposed by Lee & Wooldridge (2025). The core idea is to perform unit-specific pre-treatment transformations (like demeaning or detrending) and then collapse the panel into a cross-section for estimation.

## Key Differences

| Feature | `lwdidR` | `endid` |
|:---|:---|:---|
| **Estimation** | Ordinary Least Squares (OLS) | Engression (Neural Network) |
| **Main Target** | Average Treatment Effect (ATT) | ATT and Distributional Effects (QTE) |
| **Inference** | Analytical SEs, Wild Bootstrap, Permutation | Non-parametric Unit Bootstrap |
| **Non-linearity** | Linear only | Handles complex non-linearities |

## A Simulation Comparison

We'll simulate a panel where the treatment has a **location-scale** effect: it shifts
both the mean and the spread of the outcome distribution. This creates quantile treatment
effects (QTEs) that vary across the distribution — small or even negative at lower quantiles,
and large at upper quantiles.

The true QTE at quantile $\tau$ is:
$$\text{QTE}(\tau) = \delta + (\sigma_1 - \sigma_0) \cdot \Phi^{-1}(\tau)$$
where $\delta = 0.5$ is the location shift, $\sigma_0 = 0.5$ and $\sigma_1 = 1.5$, so
$\text{QTE}(\tau) = 0.5 + \Phi^{-1}(\tau)$. This ranges from about $-0.78$ at $\tau = 0.1$
to $1.78$ at $\tau = 0.9$.

```{r setup}
library(lwdidR)
library(endid)
set.seed(42)

# Generate synthetic panel data with location-scale treatment effect
N <- 200
T_total <- 6
tpost1 <- 4

unit <- rep(1:N, each = T_total)
time <- rep(1:T_total, times = N)
alpha <- rep(rnorm(N, sd = 0.3), each = T_total)  # Unit fixed effects
# Group 1: treated, Group 0: control
group <- rep(c(1, 0), each = N/2 * T_total)
post <- as.integer(time >= tpost1)
D <- post * (group == 1)

# Location-scale DGP: treatment shifts mean by 0.5 and increases SD from 0.5 to 1.5
epsilon <- rnorm(N * T_total)
y <- alpha + 0.5 * D + (0.5 + 1.0 * D) * epsilon

df <- data.frame(unit = unit, time = time, y = y,
                 post_treat = D, group = group)
```

### 1. Estimating with lwdidR (OLS)

`lwdidR` estimates the Average Treatment Effect using OLS.

```{r lwdid}
# Using 'post_treat' as the treatment receipt indicator
fit_lwdid <- lwdid(df, y = "y", ivar = "unit", tvar = "time", post = "post_treat",
                   vce = "hc3")
summary(fit_lwdid)
```

### 2. Estimating with endid (Engression)

`endid` uses `engression` to capture the distributional nature of the effect.

```{r endid, results='hide'}
# We use fewer epochs and bootstrap draws for speed in this vignette
# For endid, 'post' is a calendar indicator, 'dvar' is the group indicator
fit_endid <- endid(df, y = "y", ivar = "unit", tvar = "time",
                   post = "post_treat", dvar = "group",
                   num_epochs = 500, nboot = 20, silent = TRUE)
```

```{r endid_summary}
summary(fit_endid)
```

### Visualizing Quantile Treatment Effects

While `lwdidR` provides a single ATT (around 0.5, the true location shift), `endid` reveals the full picture: the treatment *hurts* at low quantiles (by increasing downside risk) and *helps substantially* at high quantiles (by increasing upside). The QTE plot should show an upward-sloping curve crossing the ATT line.

```{r plot_qte, fig.width=7, fig.height=5}
library(ggplot2)
plot(fit_endid)
```

### Adding the True QTE

We can overlay the analytical QTE to verify that `endid` recovers the true distributional effect.

```{r plot_true, fig.width=7, fig.height=5}
# True QTE: 0.5 + qnorm(tau)
taus <- seq(0.05, 0.95, by = 0.05)
true_qte <- data.frame(tau = taus, qte = 0.5 + qnorm(taus))

plot(fit_endid) +
  geom_line(data = true_qte, aes(x = tau, y = qte),
            linetype = "dashed", color = "red", linewidth = 0.8) +
  labs(caption = "Dashed red: true QTE")
```

## When to use which?

1.  **Use `lwdidR`** when you want a fast, industry-standard linear DiD estimate with well-understood frequentist properties and high-performance standard errors (like Wild Cluster Bootstrap).
2.  **Use `endid`** when you suspect the treatment effect might be non-linear, vary across the distribution, or if you want to estimate Quantile Treatment Effects (QTE). In this example, `lwdidR` correctly estimates the ATT of 0.5, but completely misses that the treatment *increases inequality* — a conclusion only visible through the distributional lens of `endid`.
