---
title: "Comparing endid and lwdidR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Comparing endid and lwdidR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Both `endid` and `lwdidR` implement the panel Difference-in-Differences (DiD) approach proposed by Lee & Wooldridge (2025). The core idea is to perform unit-specific pre-treatment transformations (like demeaning or detrending) and then collapse the panel into a cross-section for estimation.

## Key Differences

| Feature | `lwdidR` | `endid` |
|:---|:---|:---|
| **Estimation** | Ordinary Least Squares (OLS) | Engression (Neural Network) |
| **Main Target** | Average Treatment Effect (ATT) | ATT and Distributional Effects (QTE) |
| **Inference** | Analytical SEs, Wild Bootstrap, Permutation | Non-parametric Unit Bootstrap |
| **Non-linearity** | Linear only | Handles complex non-linearities |

## A Simulation Comparison

We'll simulate a simple panel where the treatment effect is non-linear—specifically, it varies by a covariate.

```{r setup}
library(lwdidR)
library(endid)
set.seed(42)

# Generate synthetic panel data
N <- 100
T_total <- 6
tpost1 <- 4

unit <- rep(1:N, each = T_total)
time <- rep(1:T_total, times = N)
X <- rep(runif(N, -1, 1), each = T_total) # Time-invariant control
# Group 1: treated, Group 0: control
group <- rep(c(1, 0), each = N/2 * T_total)
post <- as.integer(time >= tpost1)
# Treat units in group 1 after time tpost1
y <- X + (0.5 + 0.5 * X) * (post * (group == 1)) + rnorm(N * T_total, sd = 0.2)

# Create the data frame
# In lwdidR, for common-timing, 'post' should typically be the treatment-receipt indicator 
# (1 if unit i is treated at time t)
df <- data.frame(unit = unit, time = time, y = y, X = X, 
                 post_treat = post * group, group = group)
```

### 1. Estimating with lwdidR (OLS)

`lwdidR` estimates the Average Treatment Effect using OLS.

```{r lwdid}
# Using 'post_treat' as the treatment receipt indicator
fit_lwdid <- lwdid(df, y = "y", ivar = "unit", tvar = "time", post = "post_treat", 
                   controls = "X", vce = "hc3")
summary(fit_lwdid)
```

### 2. Estimating with endid (Engression)

`endid` uses `engression` to capture the distributional nature of the effect.

```{r endid, results='hide'}
# We use fewer epochs and bootstrap draws for speed in this vignette
# For endid, 'post' is a calendar indicator, 'dvar' is the group indicator
fit_endid <- endid(df, y = "y", ivar = "unit", tvar = "time", 
                   post = "post_treat", dvar = "group",
                   controls = "X",
                   num_epochs = 200, nboot = 5, silent = TRUE)
```

```{r endid_summary}
summary(fit_endid)
```

### Visualizing Quantile Treatment Effects

While `lwdidR` provides a single ATT, `endid` allows us to see how the treatment affects different parts of the distribution.

```{r plot_qte, fig.width=7, fig.height=4}
# plot(fit_endid) 
# Note: In a vignette we'd usually use ggplot2 directly or ensure it's available
library(ggplot2)
plot(fit_endid)
```

## When to use which?

1.  **Use `lwdidR`** when you want a fast, industry-standard linear DiD estimate with well-understood frequentist properties and high-performance standard errors (like Wild Cluster Bootstrap).
2.  **Use `endid`** when you suspect the treatment effect might be non-linear, vary significantly across the distribution, or if you want to estimate Quantile Treatment Effects (QTE) while still benefiting from the Lee-Wooldridge panel-to-cross-section transformation.
