---
title: "Real Data Example: Castle Doctrine"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Real Data Example: Castle Doctrine}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this vignette, we compare `lwdidR` and `endid` using the "Castle Doctrine" dataset from Cheng and Hoekstra (2013). States adopted castle doctrine laws in different years (2005--2009), making this a staggered adoption design.

## Data Preparation

We load the data and set up the treatment timing variable (`gvar`). States that never adopted the law during the sample period serve as controls.

```{r setup}
library(lwdidR)
library(endid)
library(ggplot2)
set.seed(42)

# Load the Castle Doctrine dataset
castle <- read.csv(system.file("extdata", "castle.csv", package = "lwdidR"))

# Create gvar: first treatment year (NA for never-treated)
castle$gvar <- castle$effyear
castle$gvar[is.na(castle$gvar) | castle$gvar == 0] <- NA

# Time-invariant controls
controls <- c("poverty", "unemployrt", "blackm_15_24", "whitem_15_24")

cat("States by cohort:\n")
print(table(castle$gvar[!duplicated(castle$sid)], useNA = "ifany"))
```

## Linear Estimation with lwdidR

`lwdidR` provides a robust, linear estimate of the Average Treatment Effect on the Treated (ATT) using the staggered design.

```{r lwdid}
fit_lwdid <- lwdid(
  data = castle,
  y = "lhomicide",
  ivar = "sid",
  tvar = "year",
  gvar = "gvar",
  controls = controls,
  vce = "hc3"
)
summary(fit_lwdid)
```

## Distributional Estimation with endid

`endid` uses engression to estimate how the law affects different parts of the homicide rate distribution, aggregating across all treatment cohorts.

```{r endid, results='hide'}
# Reduced epochs and bootstrap for faster rendering
fit_endid <- endid(
  data = castle,
  y = "lhomicide",
  ivar = "sid",
  tvar = "year",
  gvar = "gvar",
  controls = controls,
  rolling = "demean",
  num_epochs = 500,
  nboot = 20,
  silent = TRUE
)
```

```{r endid_summary}
summary(fit_endid)
```

## Comparing the Results

While `lwdidR` provides a single point estimate (the average effect), `endid` allows us to visualize the **Quantile Treatment Effects (QTE)**.

```{r plot_qte, fig.width=7, fig.height=4}
plot(fit_endid)
```

### Interpretation

In this example, `lwdidR` shows the average effect of castle doctrine laws on log homicide rates. The `endid` results complement this by showing how the effect varies across quantiles of the distribution. This distributional perspective can be crucial for policy analysis where the impact on the "tails" of the distribution is as important as the average impact.
